---
title: "Google Case Study Cyclistic"
author: "Zhihong Mai"
date: "2023-07-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setlocale("LC_TIME", "English")
```


### Install and loading packages
```{r warning=FALSE}
# Since we have installed these packages, we can simply load them.
library(tidyverse)
library(ggplot2)
library(lubridate)
```

### Importing the data
```{r}
# Import data from local storage. 
Jan_2022<-read.csv("C:\\Users\\maizh\\Downloads/202201-divvy-tripdata.csv",header=TRUE)
Feb_2022<-read.csv("C:\\Users\\maizh\\Downloads/202202-divvy-tripdata.csv",header=TRUE)
Mar_2022<-read.csv("C:\\Users\\maizh\\Downloads/202203-divvy-tripdata.csv",header=TRUE)
Apr_2022<-read.csv("C:\\Users\\maizh\\Downloads/202204-divvy-tripdata.csv",header=TRUE)
May_2022<-read.csv("C:\\Users\\maizh\\Downloads/202205-divvy-tripdata.csv",header=TRUE)
Jun_2022<-read.csv("C:\\Users\\maizh\\Downloads/202206-divvy-tripdata.csv",header=TRUE)
Jul_2022<-read.csv("C:\\Users\\maizh\\Downloads/202207-divvy-tripdata.csv",header=TRUE)
Aug_2022<-read.csv("C:\\Users\\maizh\\Downloads/202208-divvy-tripdata.csv",header=TRUE)
Sept_2022<-read.csv("C:\\Users\\maizh\\Downloads/202209-divvy-tripdata.csv",header=TRUE)
Oct_2022<-read.csv("C:\\Users\\maizh\\Downloads/202210-divvy-tripdata.csv",header=TRUE)
Nov_2022<-read.csv("C:\\Users\\maizh\\Downloads/202211-divvy-tripdata.csv",header=TRUE)
Dec_2022<-read.csv("C:\\Users\\maizh\\Downloads/202212-divvy-tripdata.csv",header=TRUE)
```


```{r}
# Check the column naes of each data set
colnames(Jan_2022)
colnames(Feb_2022)
colnames(Mar_2022)
colnames(Apr_2022)
colnames(May_2022)
colnames(Jun_2022)
colnames(Jul_2022)
colnames(Aug_2022)
colnames(Sept_2022)
colnames(Oct_2022)
colnames(Nov_2022)
colnames(Dec_2022)
```


```{r}
# Combine all 12 months data into a single data frame
Trips_2022<- bind_rows(Jan_2022, Feb_2022, Mar_2022, Apr_2022, May_2022, Jun_2022, Jul_2022,
                       Aug_2022, Sept_2022, Oct_2022, Nov_2022, Dec_2022)
```

## Process-Cleaning the data

```{r}
# Check the data format for each columns.
str(Trips_2022)
```

```{r}
# Converting "started_at" and "ended_at" columns to date-time forma(YYYY-MM-DD HH:MM:SS). 
Trips_2022$started_at<- ymd_hms(Trips_2022$started_at)
Trips_2022$ended_at<- ymd_hms(Trips_2022$ended_at)
str(Trips_2022)
```

```{r}
# Check for any null value
na<- any(is.na(Trips_2022))
print(na)

# Remove all the null value
Trips_2022<- Trips_2022 %>% 
              drop_na()

nna<- any(is.na(Trips_2022))
print(nna)
```

```{r}
# Check for any duplicated value
duplicated_rows<- duplicated(Trips_2022[,c("ride_id", "started_at")])
summary(duplicated_rows)
```

```{r}
# Check if each value has 16 characters in ride_id
ride_id_length<- nchar(Trips_2022$ride_id)
summary(ride_id_length)
```

```{r}
# Check the label name of "rideable_type" and "member_casual" columns.
table(Trips_2022$rideable_type)
table(Trips_2022$member_casual)
```

### Analysis and Share

```{r}
# Add columns that contains day, month, year, hours etc.
Trips_2022$Date<- as.Date(Trips_2022$started_at)
Trips_2022$Month<- format(as.Date(Trips_2022$Date), "%m")
Trips_2022$Day<- format(as.Date(Trips_2022$Date), "%d")
Trips_2022$Year<- format(as.Date(Trips_2022$Date), "%Y")
Trips_2022$Day_of_Week<- format(as.Date(Trips_2022$Date), "%A")
Trips_2022$Hours<- format(as.POSIXct(Trips_2022$started_at), "%H")
```

```{r}
# Calculate the length of each ride
Trips_2022$ride_length<- difftime(Trips_2022$ended_at, Trips_2022$started_at)
str(Trips_2022)
```

```{r}
# convert the ride_length to numeric format
Trips_2022<- Trips_2022 %>% 
  mutate(ride_length = as.numeric(as.character(ride_length)))
str(Trips_2022)
```

```{r}
# Change the value of ride_length and find out if there is any negative value 
# or values that greater than a day
Trips_2022_v2<- Trips_2022 %>% 
  filter(!(ride_length < 0 | ride_length >= 86400))
dim(Trips_2022_v2)
```

### Visulaztion
```{r}
# To prevent values in the graph from being displayed in scientific notation such as "2e+02," 
options(scipen = 100)
```

### Aggregate Rides and Trip Duration by Members and Casual Users
```{r}
# Visualize the Total Rides Taken by Members and Casual Users
Trips_2022_v2 %>% 
  group_by(member_casual) %>% 
  summarize(number_of_ride =n())

Trips_2022_v2 %>% 
  group_by(member_casual) %>% 
  summarize(number_of_ride =n()) %>% 
ggplot(mapping= aes(x= 'Content', y=number_of_ride, fill= member_casual))+  
  geom_bar(stat= 'identity', 
           position ='stack', width = 1)+            
  coord_polar(theta = 'y')+          
  labs(x='', y='', title = 'Aggregate Rides by Membership Type: Members vs Casual Users')+  
  theme_void()+    
  geom_text(aes(label = number_of_ride), position = position_stack(vjust = 0.5))

# Visualize the Trip Duration by Members and Casual Users
Trips_2022_v2 %>% 
  group_by(member_casual) %>% 
  summarize(Total_trip_duration= sum(ride_length))

Trips_2022_v2 %>% 
  group_by(member_casual) %>% 
  summarize(Total_trip_duration= sum(ride_length))  %>% 
  ggplot(mapping= aes(x= 'Content', y=Total_trip_duration, fill= member_casual))+
  geom_bar(stat= 'identity', 
           position ='stack', width = 1)+
  coord_polar(theta = 'y')+          
  labs(x='', y='', title = 'Overall Trip Duration: Members vs. Casual Users')+  
  theme_void()+    
  geom_text(aes(label = Total_trip_duration), position = position_stack(vjust = 0.5))
```

### Ride Count by Day of the Week by Members and Casual Users 

```{r}
# Comparing the Number of Rides for Casual users and Members by Day of the Week
Trips_2022_v2$Day_of_Week <- ordered(Trips_2022_v2$Day_of_Week, 
                                     levels= c("Sunday", "Monday", "Tuesday", "Wednesday",
                                               "Thursday", "Friday", "Saturday"))

Trips_2022_v2 %>% 
  group_by(member_casual, Day_of_Week) %>% 
  summarize(number_of_ride =n(), .groups = "keep") %>% 
  arrange(Day_of_Week, member_casual)

Trips_2022_v2 %>% 
  group_by(member_casual, Day_of_Week) %>% 
  summarize(number_of_ride =n(), .groups = "keep") %>% 
ggplot(mapping = aes(x=Day_of_Week, y= number_of_ride, fill=member_casual))+
  geom_bar(position = "dodge", stat="identity")+
  labs(title = "Ride Count by Day of the Week: Members vs. Casual Users", 
       x="Days of Week", y="Number of ride")
```

### Average Trip Duration by Day of the Week by Members and Casual Users 

```{r}
# Comparing the Average Ride Duration by Day of the Week by Members and Casual Users 
Trips_2022_v2 %>% 
  group_by(member_casual, Day_of_Week) %>% 
  summarize(number_of_ride =n(),
            average_duration=mean(ride_length), .groups = 'keep') %>% 
  ggplot(mapping = aes(x=Day_of_Week, y=average_duration, fill=member_casual))+
  geom_bar(position = "dodge", stat="identity")+
  labs(title = 'Average Ride Duration: Members vs. Casual Users', y= 'Average_duration')
```

```{r}
# Comparing Total Rides by Day of the Week by Members and Casual Users (Per Hour)
Trips_2022_v2 %>% 
  group_by(member_casual, Hours, Day_of_Week) %>% 
  summarize(number_of_ride =n(),
            .groups='keep') %>% 
  ggplot(mapping = aes(x=Hours, y=number_of_ride, fill=member_casual))+
  geom_bar(position = 'dodge', stat= 'identity')+
  labs(title = 'Total Rides by Day of the Week: Members vs. Casual Users (Per Hour)', y='Number of Rides')+
  facet_wrap(~Day_of_Week)
```

### Frequency of each rideable_type 
```{r}
# Compare the frequency of each rideable_type.
Trips_2022_v2 %>% 
  group_by(member_casual,rideable_type) %>% 
  summarize(number_of_ride =n(), .groups = "keep") %>% 
  ggplot(mapping = aes(x=rideable_type, y=number_of_ride, fill=member_casual))+
  geom_bar(stat = 'identity', position = 'dodge')+
  labs(title = 'Number of Rides by Rideable Type and Member vs. Causal User', x= 'Rideable Type',
       y= 'Number of Rides')
```

### Monthly Ride Count by Members and Casual Users
```{r}
# Comparing the Number of Rides for Casual Users and Members by Each Month

Trips_2022_v2 %>% 
  group_by(member_casual, Month) %>% 
  summarize(number_of_ride =n(), .groups = 'keep') %>% 
  ggplot(mapping = aes(x=Month, y=number_of_ride, fill=member_casual))+
  geom_bar(position = 'dodge', stat='identity')+
  labs(title= "Monthly Ride Count by Members and Casual Users")
```
